A Realistic Guide to Making Data Available Alongside Code to Improve Reproducibility.

### 2019-12-09 paper.Rmd
working through dotpoints and add acknowledgements outline

@@

- Explain formats like XML, as people might come into formats that wany XML.
- The power lies in the contrast to a csv data dictionary, as there is a defined, nested structure.
- Ontologies and domain specific metadata standards such as EML might also be worth mentioning.
- introduce the holy grail of the semantic web: JSON-LD

@@

- we could map out how these things link together (e.g., you can turn JSON into xml with package ...)
- make a table that describes how to do something in R, or Python, or Julia
-

@@

It is worth distinguishing between sharing data and publishing data. One can share data in numerous ways without going through the trouble of publishing it, which often requires metadata that a human must verify. Data can be shared in numerous ways, including placing it in a repository, packaging it with methods, or using various free tiers of commercial services (such as dropbox or google drive). However, one must publish data when appropriate.

@@

In **Least Moving Parts**, the data might be published with an R package, or as part of a GitHub release using piggyback @piggyback, or a in serverless databases. This approach **gets the data somewhere rather than nowhere**. Its minimal features means it is simple to maintain. In addition, piggyback also keeps the data in the same location as the code, making it easier and simpler to organise. A downsides is that it does not scale to larger data. Self hosting the data is an option, but we discourage this, as it may succumb to bit rot -.

In **Domain Specific Venue**, data can be published in journal data papers, or venues specific to the type of data. For example, an astronomy project hosts its data at the Sloan Digital Sky Survey (SDSS), and Genetic data can be hosted at GenBank (@GenBank), and so on.  Some other places to put data include the open science framework [@OSF], Dataverse, [@dataverse], and Data World, [@dataworld]. Other datasets might be produced

The purpose, use, and origin of the data is an important component to consider. Data for research has a different domain compared to data collected by governments, or by journalists. Many governments or civil organizations are now making their own data available through a government website interface. Media and journalism groups are also making their data available either through places like GitHub [@pudding-data], or organizations such as Data World, or may self host their data, such as five thirty eight [@data-538].

@@

If the dataset you are shipping has a research application, the most relevant home for it would be Zenodo, a research data repository. Launched in 2013 in a joint collaboration between openAIRE and CERN, Zenodo provides a free, archival location for any researcher to depositor their datasets. The only limits to file sizes are 50gb for individual files, which is generous enough to accommodate a large number of use cases. Zenodo is able to accommodate larger file sizes upon request.

@@

>>>>

In contrasst to a CSV data dictionary, JSON-LD has a defined, nested structure, which means more can be communicated efficiently, while still maintaining readability, and avoiding extra writing and repetition that comes with a plain CSV where everything is defined from scratch.

Another rich format is XML, the eXtensible Markup Language. This was an early iteration on the idea of a plain text format that was both human and machine readable. It is very powerful - it powers the entire Microsoft Office suite, but for metadata purposes it is not as human readable. However, one example of its usage is in EML, the Ecological Metadata Language. XML and JSON can be converted into each others respective formats

@@

- The holy grail of the semantic web: JSON-LD

@@

- we could map out how these things link together (e.g., you can turn JSON into xml with package ...)
- make a table that describes how to do something in R, or Python, or Julia

@@

It is worth distinguishing between sharing data and publishing data. Data can be shared in numerous ways without going through the trouble of publishing it, which often requires metadata that a human must verify. Data can be shared in numerous ways, including placing it in a repository, packaging it with methods, or using various free tiers of commercial services (such as dropbox or google drive). However, one must publish data when appropriate.

@@

In **Least Moving Parts**, the data might be published with an R package, or as part of a GitHub release using piggyback @piggyback, or a in serverless databases. This approach **gets the data somewhere rather than nowhere**. Its minimal features means it is simple to maintain. In addition, piggyback also keeps the data in the same location as the code, making it easier and simpler to organise. A downsides is that it does not scale to larger data. Self hosting the data is an option, but we discourage this, as it may succumb to bit rot - where data is corrupted, degraded, or servers turn off.

In **Domain Specific Venue**, data can be published in journal data papers, or venues specific to the type of data. For example, an astronomy project hosts its data at the Sloan Digital Sky Survey (SDSS), and Genetic data can be hosted at GenBank (@GenBank), and so on.

The purpose, use, and origin of the data is an important component to consider. Data for research has a different domain compared to data collected by governments, or by journalists. Many governments or civil organizations are now making their own data available through a government website interface. Media and journalism groups are also making their data available either through places like GitHub [@pudding-data], or may self host their data, such as five thirty eight [@data-538].

@@

If the dataset you are shipping has a research application, the most relevant home for it would be the research data repositories **zenodo** or **dryad**. Zenodo, Launched in 2013 in a joint collaboration between openAIRE and CERN, provides a free, archival location for any researcher to depositor their datasets. The only limits to file sizes are 50gb for individual files, which is generous enough to accommodate a large number of use cases. Zenodo is able to accommodate larger file sizes upon request. Dryad performs data curation

```{r create-draft-table, }
tibble::tribble(
  ~Organisation,              ~Pros, ~Cons,
  "FigShare",                 "",     "",
  "Dryad",                    "",     "",
  "Zenodo",                   "",     "",
  "Institutional Repository", "",     "",
) %>% 
  knitr::kable(caption = "Table Comparing common data submission platforms")
```

@@

# Acknowledgements

- Mile McBain
- Anna Krystalli
- Daniella Lowenberg
